Cosmic web alignments with the shape, angular momentum and peculiar velocities of dark matter halos

### 2014-02-18 paper/main.tex
Appendix A updated

\item
\citep{Libeskind2013} %various alignments
Studies shape and angular momentum alignments with the cosmic web
defined by the velocity shear tensor method described in this
paper.

\cite{Libeskind2013} also used the Bolshoi simulation and the halo
catalogs we use in this work. Results are reported for three mass bins $M_{\rm
  vir}<10^{11.5}$\hMsun, $11^{11.5}<M_{\rm vir}<12^{12.5}\hMsun$ and
$M_{\rm  vir}>12^{12.5}\hMsun$.   The identification of the cosmic web
is done on a grid of $256^3$ with a gaussian smoothing of $\sim
1$\hMpc over the velocity field. The way they compute this smoothed
velocity field differs from our computation. We do it based on the
momentum density field while \cite{Libesking2013} do not take into
account the mass in each cell.

The aligmnet signal for the angular momentum is very weak while the shape alignment
signal is very strong. The shape alignment is such that the eigenvector
corresponding to the smallest eigenvalue is aligned with the major
axis. This effect is stronger for more massive halos. 
In other words the major axis of a halo is aligned with a filament,
and lies on the plane that define a sheet. The angular momentum is
anti-aligned with the filament for massive halos and weakly aligned
for low mass halos.

\item
\citep{Trowland2013}

The simulation is the millennium run, which has $2160^3$
particles in a volume of $500$\hMpc on a side. This corresponds to a
particle mass of $8.6\times 10^{8}$\hMsun. The catalog uses both halos
and subhalos which were identified with SUBFIND. Only halos with more
than 500 particles were kept to get a robust computation for the
spin. For each halo the spin is defined as the sum of the angular
momentum of each particle with respect to the center of mass.

The method to define the filamentary structure is based on the
eigenvalues of the hessian of the density.However the analysis are
reported ona box of $300$\hMpc on a side. Four different gaussian
smoothing scales are used: $2.0$, $3.0$ and $5.0$\hMpc.

The authors report a slight alignment signal of spin against the
principal filament axis. By fitting the following functional form to
the $\cos(\theta)$ distribution

\begin{equation}
P(\cos\theta) =
(1-c)\sqrt{1+\frac{c}{2}}\left[1-c\left(1-\frac{3}{2}\cos^{2}\theta
  \right)\right]^{-3/2}, 
\label{eq:distro}
\end{equation}

they are able to quantify the degree of alignment ($c<0$) or
antialignment ($c>0$).  This parameterization is based on theoretical
expectactions of Tidal Torque Theory (TTT) \citep{Lee2005}. At $z=0$,
the reported value is $c = −0.035 \pm 0.004$, where the uncertainty
was calculated using bootstraping and resampling.

When the halo sample is divided between low mass and high mass halos
with a transition scale $M_{\star}=5.9\times 10^{12}$\Msun, there is
an anti-alignment above this mass and an alignment below it.

\item 
\citep{Codis2012}
Studies the alignment of the spin of dark matter halos relative to
the surrounding large scale structure and to the tidal tensor
eigenvalues.

They use a dark matter simulation with $4096^3$ DM particles in a
cubic periodic box of $2000$\hMpc on a side, which corresponds to a
particle mass of $7.7\times 10^9$\Msun. Halos are identified
using a FoF algorithm with a linking legth of 0.2 keeping all halos
with more than 40 particles, which sets the minimum halo mass to be
$3\times 10^{11}$\Msun. In their work the particles were sampled on a
$2048^3$ grid and the density field was smoothed with a gaussian
fileter over a scale of $5$\hMpc corresponding to a mass of $1.9\times
10^{14}$. The skeleton was computed over $6^{3}$ overlapping subcubes
and then reconnected.

The filament finder algorithm is based on Morse theory and defines a
Skeleton to be the set of critical lines joining the maxima of the
densit field through saddle points following the gradient. They also
compute the hessinan of the potential over the smoothed density field
to get their eigenvectors.

The spin of the halo is defined as $m_{p}\sum_{i}(r_i-\bar{r})\times
(v_i-\bar{v})$ where $\bar{r}$ is the center of mass of the halo and
$\bar{v}$ is the average velocity.

They measure the sping alignment with each one of the
eigenvectors. With repecto to the minor eigenvector $e_{1}$ there is
antialignement for masses $M>5\times10^{12}$\Msun and alignment for masses
$<5\times 10^{12}$\Msun. With respect to the intermediate eigenvector $e_2$
there is a strong alignment at high masses and no alignment for low
masses, with respecto the major eigenvector $e_{3}$ there is an
anti-alignment signal at all masses. The results from the Skeleton
algorithm are in perfect agreement with the results from the Tidal
web.  The transitional mass is weakly dependent on the smoothing scale,
varing between $1-5\times 10^{12}$\hMsun for smoothing scales between
  $1.0-5.0$\hMpc.

\item
\citep{Zhang2009}
Study the spin and shape alignment against filaments.

They use a dark matter simulation with $1024^3$ DM particles in a
periodic box of $100$ \hMpc on a side. The particle mass is
$6.92\times10^{7}$\hMsun. Dark matter haloes are found using a FOF
algorithm with a linking length of 0.2 times the interparticle
distance. Only halos with more than 500 particles are retained for
further analysis.

The angular momentum is measured with positions repect to the center
of mass and the shape is determined using the non-normalized moment of
inertia tensor.

The environment is found using the hessian of the density. The density
field was interpolated over a $1024^3$ grid and then smoothed with a
Gaussian filter of scale $R_{s} = 2.1$\hMpc. There are two methods to
define the direction of a filament. The first method uses the
eigenvalues of the hessian density, howevere they define the filament
direction with the eigenvector corresponding the single positive
eigenvalue of the hessian. The second method used a line that
connects the two terminal halos in a filament segment.

The characterization of the alingment with the $\cos(\theta)$
statistic. For the method that uses the eigenvectors, they find that
the strenght of the spin alignment decreases with halo mass. For the
shape they study the alingment of the major axis with the
filament. The find an alignment signal in all mass bins, with an stronger
effect for more massive halos.

In a final experiment they measure the spin alignment in four
different samples that separated by the strength of the shape
alignment. They find that halos anticorrelated in shape, show a strong
sping correlation going to the extreme where there is a strong spin
anticorrelation for halos with a strong shape correlation. This means
that the halos with strong spin alignment are not the same halos
showing strong shape alignment.

\item 
\citep{AragonCalvo2007} %angular momentum alignment
The method is the Multi-scale Morphology Filter which is based on the
Hessian matrix of the density field, where the density field is
computed from the particle distribution using a Delaunay tesselation
field estimatior (DTFE), which is self-adaptive. This allows them to
identify clusters, filaments and walls.

The simulation has $512^3$ particles in a cubic box of $150$\hMpc. The
mass per particle is $2\times 10^{9}$\hMsun.  Halo identitification is
done with the HOP algorithm. They keep halos with more than $50$
partices and less than $5000$, defining a mass range of$1-100\times
10^{11}$\hMsun.

The principal axes of each halo are computed from the non-normalized
inertia tensor. The inertian tensor and the angular momentum are
computed with respect to the center of mass of the halo.

They compute two angles, one with respect to the direction
defining the filaments and the other the walls. Their results make a
distinction between halos of more massive and less massive than
$10^{12}$\hMsun.

The halo spins tend to lie on the plane of the wall. This is stronger
for massive halos.The effect for filaments is weaker: low mass halos
tend to lieparallel to their host filament, while high mass halos tend
to be perpendicular.

Theres is a very strong effect for the principal axes of halos in
filaments to be strongly correlated with the direction of the
filaments. The minor axis tend to be perpendicular to the
filament. This effect is stronger for larger halos.

The effect in walls is less strong ,but still the minor axis tend to
lie perpendicular to the wall, while the other axis then to lie over
the wall. The effect is stronger for massive halos.

They find that spins and shapes of dark matter halos are significantly
correlated with each other and with the orientation of their host
structures.

\item 
\citep{Hahn2007}% various alignments with the Tweb
The method is the Tweb. They use three simulations each of $512^3$
particles, with sizes $L_1=45$\hMpc, $L_2=90$\hMpc and
$L_3=180$\hMpc, this corresponds to particle masses of $4.7, 38.0,
300\times 10^7$\hMsun. The normalization is $\sigma_8=0.9$. Halo
identification was done with a FOF algorithm with 0.2 times the
interparticle distance. They consider halos of at least 300
particles.

The web is obtained for a grid of $1024^3$ cells, the density field is
obtained with a CIC interpolation and smoothed using a Gaussian
Kernel. In the rest of the paper all the results correspond to a
smoothing scale of $R_{s}=2.1$\hMpc.

They Report on the angle between the halo
angular momentum vector and the eigenvector corresponding to
perpendicular directions to the sheets and the direction of the
filaments. This is divided in two halo populations: $5\times 10^{10} -
1.0\times 10^{12}$ and $>10^{12}$. There is a weak antialignment in
the case of the filaments and a stronger anti-alignment in the case of
the sheets. For the sheets the effect is stronger for the massive
bin. In the filaments the alignment is weak regardless of the
mass. They do not report any other significan statistic, but recognize
that they suffer from small-number statistics in voids).

They do not see any strong dependance of the environment in the
shape. They do not measure the shape alinment.

\item
\citep{Brunino2007}
Uses a DM only N-body simulation (the Millennium Run) to study the
shape and angular momentum alignment with respect to large
cosmological voids.

The voids are found with a spherical void finder. This algorithm first
looks for potential void centers in the underdense regions of the
simulation. These positions define the centers of spheres that are
empty of halos above a given mass threshold. Only voids with radius
larger than $10$\hMpc were considered for further analysis.

Halos were identified by linking particles with density exceeding 900
times the background density in order to keep the analysis of shape
and angular momentum on the core of the halo. Only halos with masses
in the range $8.6\times 10^{11}\hMsun < M_{\mathrm h}< 8.6\times
10^{12}\hMsun$ were included analized to measure their angular
momentum and the inertia tensor, both in a non-normalized way.

For the purpose of measuring alignments a second cut is imposed
whereby only halos within a shell of radius $R_{\mathrm void}<r<1.05R_{\mathrm
void}$. The alignment is then measured between the vector connecting
the center of the void to the halo in question and the vector of
interest.

The results show that there is a strong alignment of halo shape in
such a way that the major axis is parallel to the shell surrounding
the void. In the case of angular momentum there is not a significant
signal for alignment.

\item
\citep{Basilakos2006}
Use a cosmological SPH+N-body simulation to measure the alignment of
cluster halos with their parent supercluster. For both the cluster
halos and parent super-cluster they define the shape via the
non-normalized inertia tensor.  The find that strenght of the
alignments increases with the degree of filamentarity of the
supercluster.

The simulation has $2\times 512^3$ particles in a box of side
$500$\hMpc. The dark matter particle mass is $6.6\times
10^{10}$\hMsun while for SPH particles is $1.2\times
10^{10}$\hMsun. The halo finding is done with a FOF algorithm with a
linking length of 0.17 and keep objects with more than 100 partcles.

\item
\citep{Lee2002}%Obs. spin alignment
Observational measurement for the alignment of galaxy spin axes with
the local tidal shear field. For the measurement of shear, we have used
the Point Source Catalog Redshift (PSCz) survey (a complete redshift survey
from the IRAS Point Source Catalog) data. This was done down to a
radial comoving distance of $\sim 150$\hMpc.
\item
\citep{Hatton2001} %angular momentum alignment
DM matter only simulation. $256^3$ particles, $100$\hMpc, particle
mass. Look for alignment of angular momentum vectors with neighbohring
structures or other massive structures. They match halos using close
pair statistics. They do not find any evidence for a statistically
significant mutual alignment of halos on any escale. 
\end{itemize}

>>>>

\item
\cite{Libeskind2013} %various alignments

They study the shape and angular momentum alignments with the cosmic
web defined by the velocity shear tensor method described in this
paper.  \cite{Libeskind2013} used the Bolshoi simulation and the halo
catalogs we use in this work. Results are reported for three mass bins $M_{\rm
  vir}<10^{11.5}$\hMsun, $11^{11.5}<M_{\rm vir}<12^{12.5}\hMsun$ and
$M_{\rm  vir}>12^{12.5}\hMsun$. The identification of the cosmic web
is done on a grid of $256^3$ with a gaussian smoothing of $\sim
1$\hMpc over the velocity field. The way they compute this smoothed
velocity field differs from our computation. We do it based on the
momentum density field while \cite{Libesking2013} do not take into
account the mass in each cell.

The aligmnet signal for the angular momentum is weak while the shape
alignmentsignal is very strong. The shape alignment is such that the
eigenvector corresponding to the smallest eigenvalue is aligned with
the major axis. This effect is stronger for more massive halos.  In
other words the major axis of a halo is aligned with a filament, and
lies on the plane that define a sheet. The angular momentum is
anti-aligned with the filament for massive halos and weakly aligned
for low mass halos.

\item
\cite{Trowland2013}

They used the Millennium Run, which has $2160^3$ particles in a volume
of $500$\hMpc on a side. This corresponds to a particle mass of
$8.6\times 10^{8}$\hMsun. The catalog uses both halos and subhalos
identified with SUBFIND. Only halos with more than 500 particles were
kept to get a robust computation for the spin. The angular momentum is
defined as the sum of the angular momentum of each particle with 
respect to the center of mass.

The method to define the filamentary structure is based on the
eigenvalues of the hessian of the density.  However, the analysis is
performed on a box of $300$\hMpc on a side. Four different gaussian
smoothing scales are used: $2.0$, $3.0$ and $5.0$\hMpc.

By fitting the following functional form tothe $\cos(\theta)$ distribution

\begin{equation}
P(\cos\theta) =
(1-c)\sqrt{1+\frac{c}{2}}\left[1-c\left(1-\frac{3}{2}\cos^{2}\theta
  \right)\right]^{-3/2}, 
\label{eq:distro}
\end{equation}
%
they are able to quantify the degree of alignment ($c<0$) or
antialignment ($c>0$).  This parameterization is based on theoretical
expectactions of Tidal Torque Theory (TTT) \citep{Lee2005}. At $z=0$,
the reported value is $c = −0.035 \pm 0.004$, where the uncertainty
was calculated using bootstraping and resampling.

When the halo sample is divided between low mass and high mass halos
with a transition scale $M_{\star}=5.9\times 10^{12}$\Msun, there is
a weak alignment signal of the angular momentum against the
principal filament axis for halos above that mass, for halos below
that scale there is a weak anti-alignment.

\item 
\cite{Codis2012}

They studies the alignment of the angular momentum dark relative to
the surrounding large scale structure and to the tidal tensor
eigenvalues.

They use a dark matter simulation with $4096^3$ DM particles in a
cubic periodic box of $2000$\hMpc on a side, which corresponds to a
particle mass of $7.7\times 10^9$\Msun. Halos are identified
using a FoF algorithm with a linking length of 0.2 keeping all halos
with more than 40 particles, which sets the minimum halo mass to be
$3\times 10^{11}$\Msun. In their work the particles were sampled on a
$2048^3$ grid and the density field was smoothed with a gaussian
fileter over a scale of $5$\hMpc corresponding to a mass of $1.9\times
10^{14}$. The skeleton was computed over $6^{3}$ overlapping subcubes
and then reconnected.

The filament finder algorithm is based on Morse theory and defines a
Skeleton to be the set of critical lines joining the maxima of the
density field through saddle points following the gradient. They also
compute the hessian of the potential over the smoothed density field
to get their eigenvectors.

The angular momentum of the halo is defined as $m_{p}\sum_{i}(r_i-\bar{r})\times
(v_i-\bar{v})$ where $\bar{r}$ is the center of mass of the halo and
$\bar{v}$ is the average velocity.

They measure the alignment with each one of the eigenvectors. With
repecto to the minor eigenvector $\vec{e}_{3}$  (the filament direction) there is anti-alignement for masses $M>5\times10^{12}$\Msun and alignment for masses
$<5\times 10^{12}$\Msun; with respect to the intermediate eigenvector $\vec{e}_2$
there is a strong alignment at high masses and no alignment for low
masses; with respect to the major eigenvector $\vec{e}_{1}$
(normal to the wall plane) there is an anti-alignment signal at all
masses. The results from the Skeleton algorithm are in
agreement with the results from the Tidal web.  The transitional mass
is weakly dependent on the smoothing scale, varing between $1-5\times
10^{12}$\hMsun for smoothing scales between $1.0-5.0$\hMpc.

\item
\citep{Zhang2009}

They studied the angular momentum and shape alignment against
filaments.  They used a dark matter simulation with $1024^3$ DM particles in a
periodic box of $100$ \hMpc on a side. The particle mass is
$6.92\times10^{7}$\hMsun. Dark matter haloes are found using a FOF
algorithm with a linking length of 0.2 times the interparticle
distance. Only halos with more than 500 particles are retained for
further analysis. The angular momentum is measured with positions
repect to the center of mass and the shape is determined using the
non-normalized moment of inertia tensor.

The environment is found using the hessian of the density. The density
field was interpolated over a $1024^3$ grid and then smoothed with a
Gaussian filter of scale $R_{s} = 2.1$\hMpc. There are two methods to
define the direction of a filament. The first method uses the
eigenvalues of the hessian density; they take the filament
direction to be the eigenvector corresponding the single positive
eigenvalue of the hessian. The second method used a line that
connects the two terminal halos in a filament segment.

For the method that uses the eigenvectors, they find that the strenght
of the angular momentum alignment decreases with halo mass. For the
shape they study the alingment of the major axis with the
filament. The find an alignment signal in all mass bins, with an
stronger effect for more massive halos.

In a final experiment they measure the angular momentum alignment in four
different samples split by the strength of the shape alignment. They
find that halos anti-aligned in shape, show a strong angular momentum
correlation; and a strong angular momentum anti-alignment for halos
with a strong shape alignment.

\item 
\citep{AragonCalvo2007} %angular momentum alignment

The used the Multi-Scale Morphology Filter to describe th filamentary
structure. The method is based on the Hessian matrix of the density
field, which is computed from the particle distribution using a
Delaunay tesselation field estimatior (DTFE). This allows them to
identify clusters, filaments and walls.

They used a simulation with $512^3$ particles in a cubic box of $150$\hMpc. The
mass per particle is $2\times 10^{9}$\hMsun.  Halo identitification is
done with the HOP algorithm. They keep halos with more than $50$
particles and less than $5000$, defining a mass range of$1-100\times
10^{11}$\hMsun. The principal axes of each halo are computed from the
non-normalized inertia tensor. The inertia tensor and the angular
momentum are computed with respect to the center of mass of the halo.

They compute two angles, one with respect to the direction
defining the filaments and the other the walls. Their results make a
distinction between halos of more massive and less massive than
$10^{12}$\hMsun. The angular momentum tends to lie along the plane of
the wall, with a stronger alignment for massive halos. The effect for
filaments is weaker, low mass halos tend to alignt along the filament,
while high mass halos tend to be anti-aligned.

For the shape they find a very strong alignment along filaments. In
walls the major axis lies along the wall. Both alignments are stronger
for massive halos.

\item

\citep{Hahn2007}% various alignments with the Tweb

The used the Tweb method applied on three simulations each of $512^3$
particles, with sizes $L_1=45$\hMpc, $L_2=90$\hMpc and
$L_3=180$\hMpc, this corresponds to particle masses of $4.7, 38.0,
300\times 10^7$\hMsun. Halo identification was done with a FOF
algorithm with 0.2 times the interparticle distance. They considered
halos of at least 300  particles.

The web is obtained for a grid of $1024^3$ cells, the density field is
obtained with a CIC interpolation and smoothed using a Gaussian
Kernel. All the results correspond to a smoothing scale of
$R_{s}=2.1$\hMpc.

They report on the angle between the halo angular momentum vector and
the eigenvector corresponding to perpendicular directions to the
sheets and the direction of the filaments. This is divided into two halo
populations according to mass; low mass $5\times 10^{10} - 1.0\times 10^{12}$ and
high mass $>10^{12}$. They find a weake anti-alignment for filaments
and a stronger anti-alignment in the case of the sheets. For the
sheets the effect is stronger for the massive bin. The anti-alignment
along filaments is weak regardless of the mass. They do not report any
other significan statistic, but recognize that they suffer from
small-number statistics in voids.

\end{itemize}
